Geometry of Ranked Nearest Neighbour Interchange Space of Phylogenetic Trees

### 2020-07-03 rnni_convexity.tex
mark what has been moved

\newcommand{\np}{\mathcal{NP}}
\newcommand{\parent}{\mathrm{parent}}
\newcommand{\mrca}{\mathrm{mrca}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\nni}{\mathrm{NNI}}
\newcommand{\rnni}{\mathrm{RNNI}}
\newcommand{\rnniu}{\mathrm{RNNIu}}
\newcommand{\tbr}{\mathrm{TBR}}
\newcommand{\spr}{\mathrm{SPR}}
\newcommand{\csort}{\textsc{Caterpillar Sort}}
\newcommand{\findpath}{\textsc{FindPath}}
\newcommand{\mdtree}{\textsc{MDTree}}
\newcommand{\sidma}[1]{\todo[color=green]{#1}}

@@

\sidma{Since this is the crux of the paper, expand, add images, and work through an example of the algorithm}
Let $T$ and $R$ be arbitrary trees and $[C_1, \ldots, C_{n-2}]$ be the cluster representation of $R$ (the ``destination'' tree).
Then $C_i$ is induced by the node of rank $i$ for $i = 1, \ldots, n-2$.
$\findpath$ computes a path by iteratively extending a sequence $p$ of trees starting from $T$.
The algorithm terminates when $p$ is an $\rnni$ path from $T$ to $R$.
At step $k = 1, \ldots, n-2$ of $\findpath$ we consider the cluster $C_k$ and extend path $p$ as follows.
We search for a node in the last tree $T'$ of $p$ from the previous step that induces the smallest cluster containing all elements of $C_k$.
This node is denoted by $\mrca_{T'}(C_k)$ (short for \emph{most recent common ancestor}).
We then extend $p$ by adding the tree obtained from $T'$ by performing an $\rnni$ move that decreases the rank of $\mrca_{T'}(C_k)$.
We continue extending $p$ in this way by repeatedly performing $\rnni$ moves until the rank of $\mrca_{T'}(C_k) = k$, thus completing step $k$.
\sidma{Reword [line 5 of algorithm] and/or expand discussion in text to demonstrate what this is doing}

@@

\sidma{While is highly likely that the worst-case complexity is quadratic, you need to demonstrate why that is the case. The simplest, and most convincing way would be to write this as a proposition and then go line-by-line detailing the running time.}
Clearly, the worst-case complexity of $\findpath$ is quadratic in the number of taxa.

>>>>

\newcommand{\np}{\mathcal{NP}}
\newcommand{\parent}{\mathrm{parent}}
\newcommand{\mrca}{\mathrm{mrca}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\nni}{\mathrm{NNI}}
\newcommand{\rnni}{\mathrm{RNNI}}
\newcommand{\rnniu}{\mathrm{RNNIu}}
\newcommand{\tbr}{\mathrm{TBR}}
\newcommand{\spr}{\mathrm{SPR}}
\newcommand{\csort}{\textsc{Caterpillar Sort}}
\newcommand{\findpath}{\textsc{FindPath}}
\newcommand{\mdtree}{\textsc{MDTree}}
\newcommand{\sidma}[1]{\todo[color=green]{#1}}
\newcommand{\moved}[1]{\todo[color=red]{start moved bock}{#1}\todo[color=red]{end moved block}}

@@

\moved{
\sidma{Since this is the crux of the paper, expand, add images, and work through an example of the algorithm}
Let $T$ and $R$ be arbitrary trees and $[C_1, \ldots, C_{n-2}]$ be the cluster representation of $R$ (the ``destination'' tree).
Then $C_i$ is induced by the node of rank $i$ for $i = 1, \ldots, n-2$.
$\findpath$ computes a path by iteratively extending a sequence $p$ of trees starting from $T$.
The algorithm terminates when $p$ is an $\rnni$ path from $T$ to $R$.
At step $k = 1, \ldots, n-2$ of $\findpath$ we consider the cluster $C_k$ and extend path $p$ as follows.
We search for a node in the last tree $T'$ of $p$ from the previous step that induces the smallest cluster containing all elements of $C_k$.
This node is denoted by $\mrca_{T'}(C_k)$ (short for \emph{most recent common ancestor}).
We then extend $p$ by adding the tree obtained from $T'$ by performing an $\rnni$ move that decreases the rank of $\mrca_{T'}(C_k)$.
We continue extending $p$ in this way by repeatedly performing $\rnni$ moves until the rank of $\mrca_{T'}(C_k) = k$, thus completing step $k$.
\sidma{Reword [line 5 of algorithm] and/or expand discussion in text to demonstrate what this is doing}

@@

\sidma{While is highly likely that the worst-case complexity is quadratic, you need to demonstrate why that is the case. The simplest, and most convincing way would be to write this as a proposition and then go line-by-line detailing the running time.}
Clearly, the worst-case complexity of $\findpath$ is quadratic in the number of taxa.}
